---
title: "Regression"
author: "Sai Pramod Bhupathi"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{R}
library(dplyr)
library(ggplot2)
library(readxl)

data <- read.csv("E:/GSU_Study/COURSE_SCHEDULE/FALL/MSA8190-Statistical Foundation/Practise HW/Bias_correction_ucl.csv")


missing_values <- sapply(data, function(x) sum(is.na(x)))

missing_data_frame <- data.frame(variable = names(missing_values), missing_count = missing_values)

missing_data_frame <- filter(missing_data_frame, missing_count > 0)

# Plotting the missing values
ggplot(missing_data_frame, aes(x = variable, y = missing_count)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "Variables", y = "Number of Missing Values", title = "Missing Values in Each Variable")

```

```{r}

missing_data_frame <- arrange(missing_data_frame, desc(missing_count))

print(missing_data_frame)

# 1% of the data is missing
```

```{r}
#removing missing values

data_clean =na.omit(data)

data_clean <- select(data_clean,- station)




```

```{r}
str(data_clean)
```

```{r}
data_sorted <- data_clean %>% arrange(as.Date(Date, format="%Y-%m-%d"))

n <- nrow(data_sorted)
train_index <- floor(0.6 * n)
valid_index <- floor(0.8 * n)

```

```{r}
# Split the data into Train, Valid, and Test sets
Train <- data_sorted[1:train_index, ]
Valid <- data_sorted[(train_index + 1):valid_index, ]
Test <- data_sorted[(valid_index + 1):n, ]

```

```{r}
#cross checking for null values
sapply(Train, function(x) sum(is.na(x)))
sapply(Valid, function(x) sum(is.na(x)))
sapply(Test, function(x) sum(is.na(x)))


```

```{r}

model_initial <- lm(Next_Tmax ~ Present_Tmax+Present_Tmin+LDAPS_RHmin+LDAPS_RHmax+LDAPS_Tmax_lapse+
                      LDAPS_Tmin_lapse+LDAPS_WS+LDAPS_LH+LDAPS_CC1+LDAPS_CC2+LDAPS_CC3+LDAPS_CC4+
                      LDAPS_PPT1+LDAPS_PPT2+LDAPS_PPT3+LDAPS_PPT4+lat+lon+DEM+Slope+Solar.radiation, data = Train) # Excluding Date from 

predictions_valid <- predict(model_initial, Valid)

# Calculate RMSE on the validation set
rmse_valid <- sqrt(mean((Valid$Next_Tmax - predictions_valid)^2))

summary(model_initial)

# Plotting residuals to check for patterns
residuals <- resid(model_initial)
ggplot(Train, aes(x = fitted(model_initial), y = residuals)) + 
    geom_point() + 
    geom_hline(yintercept = 0, linetype = "dashed") + 
    xlab("Fitted Values") + ylab("Residuals") + 
    ggtitle("Residuals vs Fitted")

# Output RMSE
rmse_valid

library(lmtest)
library(zoo)
shapiro.test(model_initial$residuals)
bptest(model_initial)

```

Multiple R-squared (0.738)

Adjusted R-squared (0.737): very close to the Multiple R-squared. This tells us that 73% of variation is being explained in the model for the predicted variable

F-statistic (609.7): This is high, and the extremely low p-value (< 2.2e-16) confirms that the model is statistically significant.

Significance of Variables

Highly Significant Variables (p < 0.001): Includes variables like Present_Tmax, LDAPS_RHmin, LDAPS_RHmax, LDAPS_Tmax_lapse, LDAPS_WS, LDAPS_LH, LDAPS_CC1, LDAPS_CC3, LDAPS_CC4, LDAPS_PPT2, lat, lon, Slope, DEM

Moderately Significant Variables: Such as Present_Tmin, LDAPS_Tmin_lapse, LDAPS_CC2, LDAPS_PPT1, LDAPS_PPT4, Solar.radiation.

Insignificant Variables: LDAPS_PPT3,LDAPS_PPT4 has a high p-value, suggesting it does not significantly contribute to the model.

Residuals
The summary of residuals shows a range from -5.7 to 6.19 with quartiles spread around zero, indicating a reasonably symmetric distribution of residuals.

From the Sharipo test we conclude that the residuals dont follow normal distribution as pvalue is lesser than 0.05.


```{R}

#REMOVING LDAPS_PPT3
model_initial1 <- lm(Next_Tmax ~ Present_Tmax+Present_Tmin+LDAPS_RHmin+LDAPS_RHmax+LDAPS_Tmax_lapse+
                      LDAPS_Tmin_lapse+LDAPS_WS+LDAPS_LH+LDAPS_CC1+LDAPS_CC2+LDAPS_CC3+LDAPS_CC4+
                      LDAPS_PPT1+LDAPS_PPT2+LDAPS_PPT4+lat+lon+DEM+Slope+Solar.radiation, data = Train) 

predictions_valid1 <- predict(model_initial1, Valid)

rmse_valid1 <- sqrt(mean((Valid$Next_Tmax - predictions_valid1)^2))

summary(model_initial1)


rmse_valid1
```
```{R}
#REMOVING LDAPS_PPT4
model_initial1 <- lm(Next_Tmax ~ Present_Tmax+Present_Tmin+LDAPS_RHmin+LDAPS_RHmax+LDAPS_Tmax_lapse+
                      LDAPS_Tmin_lapse+LDAPS_WS+LDAPS_LH+LDAPS_CC1+LDAPS_CC2+LDAPS_CC3+LDAPS_CC4+
                      LDAPS_PPT1+LDAPS_PPT2+lat+lon+DEM+Slope+Solar.radiation, data = Train) 

predictions_valid1 <- predict(model_initial1, Valid)

rmse_valid1 <- sqrt(mean((Valid$Next_Tmax - predictions_valid1)^2))

summary(model_initial1)


rmse_valid1

```

```{r}
# Create a new feature: Daily Temperature Range
Train$Temp_Range <- Train$Present_Tmax - Train$Present_Tmin
Valid$Temp_Range <- Valid$Present_Tmax - Valid$Present_Tmin



# Update the model with the new feature
model_with_feature <- lm(Next_Tmax ~  LDAPS_RHmin + LDAPS_RHmax +
                         LDAPS_Tmax_lapse + LDAPS_Tmin_lapse + LDAPS_WS + LDAPS_LH + LDAPS_CC1 +
                         LDAPS_CC2 + LDAPS_CC3 + LDAPS_CC4 + LDAPS_PPT1 + LDAPS_PPT2 + 
                          lat + lon + DEM + Slope + Solar.radiation + Temp_Range, 
                         data = Train)

predictions_valid_with_feature <- predict(model_with_feature, Valid)

rmse_valid_with_feature <- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature)^2))

summary(model_with_feature)

rmse_valid_with_feature

```

```{R}
#removing LDAPS_CC2 ,Solar radiation
#droping the model due to poor performance

model_with_feature1a <- lm(Next_Tmax ~  LDAPS_RHmin + LDAPS_RHmax +
                         LDAPS_Tmax_lapse + LDAPS_Tmin_lapse + LDAPS_WS + LDAPS_LH + LDAPS_CC1  + LDAPS_CC3 + LDAPS_CC4 + LDAPS_PPT1 + LDAPS_PPT2  +
                          lat + lon + DEM + Slope +  Temp_Range, 
                         data = Train)

predictions_valid_with_feature1a <- predict(model_with_feature1a, Valid)

rmse_valid_with_feature1a <- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature1a)^2))

summary(model_with_feature1a)

rmse_valid_with_feature1a

```


```{r}

Train$LDAPS_Range <- Train$LDAPS_RHmax - Train$LDAPS_RHmin
Valid$LDAPS_Range <- Valid$LDAPS_RHmax - Valid$LDAPS_RHmin

model_with_feature2 <- lm(Next_Tmax ~ Present_Tmax + Present_Tmin  +
                         LDAPS_Tmax_lapse + LDAPS_Tmin_lapse + LDAPS_WS + LDAPS_LH + LDAPS_CC1 +
                         LDAPS_CC2 + LDAPS_CC3 + LDAPS_CC4 + LDAPS_PPT1 + LDAPS_PPT2  + lat + lon + DEM + Slope + Solar.radiation +LDAPS_Range, 
                         data = Train)

predictions_valid_with_feature2 <- predict(model_with_feature2, Valid)

rmse_valid_with_feature2 <- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature2)^2))

summary(model_with_feature2)

rmse_valid_with_feature2

```
After removing the insignificant variables model performance is same as initial model

```{R}

Train$Interaction_WS_LH <- Train$LDAPS_WS * Train$LDAPS_LH
Valid$Interaction_WS_LH <- Valid$LDAPS_WS * Valid$LDAPS_LH

# Update the model with the interaction term
model_with_feature3 <- lm(Next_Tmax ~ Present_Tmax + Present_Tmin + LDAPS_RHmin + 
    LDAPS_RHmax + LDAPS_Tmax_lapse + LDAPS_Tmin_lapse +  LDAPS_CC1 + LDAPS_CC2 + LDAPS_CC3 + LDAPS_CC4 + 
    LDAPS_PPT1 + LDAPS_PPT2 + lat + lon + DEM + Slope + Solar.radiation + Interaction_WS_LH, data = Train)


predictions_valid_with_feature3 <- predict(model_with_feature3, Valid)

rmse_valid_with_feature3 <- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature3)^2))

summary(model_with_feature3)

rmse_valid_with_feature3
```



```{R}
Train$Present_Tmax_squared <- Train$Present_Tmax^2
Valid$Present_Tmax_squared <- Valid$Present_Tmax^2


# Update the model with the squared term
model_with_feature4 <- lm(Next_Tmax ~  Present_Tmax_squared + 
    Present_Tmin + LDAPS_RHmin + LDAPS_RHmax + LDAPS_Tmax_lapse + 
    LDAPS_Tmin_lapse + LDAPS_WS + LDAPS_LH + LDAPS_CC1 + LDAPS_CC2 + 
    LDAPS_CC3 + LDAPS_CC4 + LDAPS_PPT1 + LDAPS_PPT2 + lat + lon +
    DEM + Slope + Solar.radiation, data = Train)


predictions_valid_with_feature4 <- predict(model_with_feature4, Valid)

rmse_valid_with_feature4 <- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature4)^2))

summary(model_with_feature4)

rmse_valid_with_feature4

```

```{R}
# Log-transform LDAPS_WS
Train$LDAPS_WS_log <- log(Train$LDAPS_WS + 1)  # Adding 1 to avoid log(0)
Valid$LDAPS_WS_log <- log(Valid$LDAPS_WS + 1)

# Update the model with the transformed variable
model_with_feature5 <- lm(Next_Tmax ~ Present_Tmax + Present_Tmin + LDAPS_RHmin + 
    LDAPS_RHmax + LDAPS_Tmax_lapse + LDAPS_Tmin_lapse + LDAPS_LH + 
    LDAPS_CC1 + LDAPS_CC2 + LDAPS_CC3 + LDAPS_CC4 + LDAPS_PPT1 + LDAPS_PPT2 +
    lat + lon + DEM + Slope + Solar.radiation + LDAPS_WS_log, data = Train)


predictions_valid_with_feature5 <- predict(model_with_feature5, Valid)

rmse_valid_with_feature5<- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature5)^2))

summary(model_with_feature5)

rmse_valid_with_feature5

```

```{R}

Train$LDAPS_Total <- Train$LDAPS_CC1+Train$LDAPS_CC2+Train$LDAPS_CC3+Train$LDAPS_CC4
Valid$LDAPS_Total <- Valid$LDAPS_CC1+Valid$LDAPS_CC2+Valid$LDAPS_CC3+Valid$LDAPS_CC4

# Update the model with the transformed variable
model_with_feature6 <- lm(Next_Tmax ~ Present_Tmax + Present_Tmin + LDAPS_RHmin + 
    LDAPS_RHmax + LDAPS_Tmax_lapse + LDAPS_Tmin_lapse + LDAPS_LH + 
     + LDAPS_PPT1 + LDAPS_PPT2 +LDAPS_WS_log+
    lat + lon + DEM + Slope + Solar.radiation + LDAPS_Total, data = Train)


predictions_valid_with_feature6 <- predict(model_with_feature6, Valid)

rmse_valid_with_feature6<- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature6)^2))

summary(model_with_feature6)

rmse_valid_with_feature6



```

```{r}

# Update the model with removing solar radiation
model_with_feature6 <- lm(Next_Tmax ~ Present_Tmax + Present_Tmin + LDAPS_RHmin + 
    LDAPS_RHmax + LDAPS_Tmax_lapse + LDAPS_Tmin_lapse + LDAPS_LH + 
     + LDAPS_PPT1 + LDAPS_PPT2 +LDAPS_WS_log+
    lat + lon  + Slope + DEM + LDAPS_Total, data = Train)


predictions_valid_with_feature6 <- predict(model_with_feature6, Valid)

rmse_valid_with_feature6<- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature6)^2))

summary(model_with_feature6)

rmse_valid_with_feature6



```

```{r}


Train$LDAPS_PPT_TOTAL <- Train$LDAPS_PPT2+Train$LDAPS_PPT1
Valid$LDAPS_PPT_TOTAL <- Valid$LDAPS_PPT2+Valid$LDAPS_PPT1

model_with_feature6a <- lm(Next_Tmax ~ Present_Tmax + Present_Tmin + LDAPS_RHmin + 
    LDAPS_RHmax + LDAPS_Tmax_lapse + LDAPS_Tmin_lapse + LDAPS_LH +
     + LDAPS_PPT_TOTAL+LDAPS_WS_log+DEM+
    lat + lon  + Slope +  LDAPS_Total, data = Train)


predictions_valid_with_feature6a <- predict(model_with_feature6a, Valid)

rmse_valid_with_feature6a<- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature6a)^2))

summary(model_with_feature6a)

rmse_valid_with_feature6a


```
```{r}

bestmodel <- lm(Next_Tmax ~ Present_Tmax +  LDAPS_RHmin + 
    LDAPS_RHmax + LDAPS_Tmax_lapse + LDAPS_Tmin_lapse + LDAPS_LH +
     + LDAPS_PPT_TOTAL+LDAPS_WS+DEM+
    lat + lon  + Slope +  LDAPS_Total, data = Train)


predictions_valid_with_feature6a <- predict(bestmodel, Valid)

rmse_valid_with_feature6a<- sqrt(mean((Valid$Next_Tmax - predictions_valid_with_feature6a)^2))

summary(bestmodel)
rmse_valid_with_feature6a

```


Multiple R-squared (0.729): This value indicates that approximately 72% of the variability in Next_Tmax is explained by the model.

Adjusted R-squared (0.728): Very close to the Multiple R-squared, suggesting that the number of predictors is appropriate for the model.

F-statistic (941.1): This is very high, and the associated p-value is significantly low (< 2.2e-16), indicating that the overall model is statistically significant.

Significance of Variables

All the variables are highly significant LDAPS_Tmin_lapse in predicting the Next_Tmax.

LDAPS_Tmin_lapse is also significant but lesser compared to other variables


Residuals Distribution: The range of residuals is from -5.6 to 5.6, with the quartiles suggesting that the majority of residuals are clustered closer to zero, which is a good sign.

```{r}


# Plotting residuals to check for patterns
residuals <- bestmodel$residuals
ggplot(Train, aes(x = fitted(bestmodel), y = residuals)) + 
    geom_point() + 
    geom_hline(yintercept = 0, linetype = "dashed") + 
    xlab("Fitted Values") + ylab("Residuals") + 
    ggtitle("Residuals vs Fitted")

shapiro.test(residuals(bestmodel))
bptest(bestmodel)


```

```{r}
# Initial Model
# Make predictions on the test set
predictions_test_initial <- predict(model_initial, Test)

# Calculate RMSE on the test set for the initial model
rmse_test_initial <- sqrt(mean((Test$Next_Tmax - predictions_test_initial)^2))

#feature engineering variables from the best model
Test$LDAPS_PPT_TOTAL <- Test$LDAPS_PPT2+Test$LDAPS_PPT1
Test$LDAPS_Total <- Test$LDAPS_CC1+Test$LDAPS_CC2+Test$LDAPS_CC3+Test$LDAPS_CC4


# Improved Model
# Make predictions on the test set for the model with the new feature
predictions_test_with_feature6 <- predict(bestmodel, Test)

# Calculate RMSE on the test set for the improved model
rmse_test_with_feature6 <- sqrt(mean((Test$Next_Tmax - predictions_test_with_feature6)^2))


# Print RMSE for the initial and improved models on the test set
print("RMSETest for Initial Model:")
summary(predictions_test_initial)
print(rmse_test_initial)

print("RMSETest for Best Model")
summary(predictions_test_with_feature6)
print(rmse_test_with_feature6)
```